---
title: "A Paradigm Gap in Turkish - Corpus"
author: "Muhammed İleri"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/Desktop/paradigm gap data/corpus/materials')
theme_set(theme_light())
```

```{r library, include=FALSE}
library(tidyverse)
library(magrittr)
```

## Corpus Data

This file is a supplementary to the project **A Paradigm Gap in Turkish**. In what follows, we are plotting the frequency distributions of agreement markers depending on a given stem based on corpus data that we got from TrTenTen (via SketchEngine.eu) and TSCorpus.

### istiyor+AGR(i)

In this one we will plot the frequencies of the agreement suffixes co-occuring with the stem 'want.IMPF'.

```{r istiyor+AGR, echo = T, results = 'hide'}
person <- factor(c("1sg", "1pl", "2sg", "2pl", "3sg", "3pl"), levels = c("1sg", "1pl", "2sg", "2pl", "3sg", "3pl"))

istemek_counts <- c(744095, 365941, 26847, 52413, 346793, 91838)

istiyor_df <- tibble(person, istemek_counts)

istiyor_plot <- 
  istiyor_df %>% 
  mutate(props = istemek_counts/sum(istemek_counts)) %>%
  ggplot(aes(person, props, fill = person)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels=scales::percent) +
  xlab("person-number") +
  ylab("Relative Frequency") +
  ggtitle("istiyor + AGR      'want.IMPF' ") +
  guides(fill="none")

istiyor_plot
```

### V+AsI+AGR(i)

The plots in the grid in this part represents the frequency distribution of agreement markers attached to the verbal roots al 'buy', git 'go', gör 'see', yap 'do'.

```{r pressure, echo = T, results = 'hide'}
top4_desiderative <- read_csv("top4verbs_desiderative_freqs.csv")

relative_frequency_table <- 
  top4_desiderative %>% 
  group_by(person) %>%
  summarize(n_occurence = sum(count)) %>% 
  mutate(relative_frequencies = (n_occurence/sum(n_occurence)))

V_AsI_AGR <- 
  top4_desiderative %>%
  mutate(person = factor(person, levels = c("1sg", "1pl", "2sg", "2pl", "3sg", "3pl")),
         verb = factor(verb)) %>%
  group_by(verb) %>%
  mutate(proportion = count/sum(count)) %>%
  ggplot(aes(person, proportion)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(~verb) +
  ylab("Relative Frequency") +
  xlab("person-number") +
  guides(fill="none")
```

### Possessives

Template-final counts of possessive agreement suffixes taken from the TS Corpus.

```{r possessives, echo = T, results = 'hide'}
possessives <- c("P1sg", "P1pl", "P2sg", "P2pl", "P3sg", "P3pl")

template_final_counts <- c(969591, 1012695, 126734, 454622, 29565419, 652394)

template_final_poss_freq <- tibble(possessives, template_final_counts)

template_final_poss_relative_freq_plot <- 
  template_final_poss_freq %>%
  mutate(possessives = factor(possessives, levels = c("P1sg", "P1pl", "P2sg", "P2pl", "P3sg", "P3pl"))) %>%
  mutate(relative_freqs = template_final_counts/sum(template_final_counts)) %>%
  ggplot(aes(possessives, relative_freqs, fill = possessives)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks=seq(0.1, 0.9, by=0.2), labels=scales::percent) +
  ggtitle("Relative Word-Final Frequencies of Possessive Suffixes") +
  ylab("Relative Frequency") +
  xlab("Possessive Suffixes") +
  guides(fill="none")

template_final_poss_relative_freq_plot

```

### -mA Nominalization

The plot below illustrates the relative frequency distribution of possessive agreement suffixes when given a stem endign with the nominalizer suffix -mA. The frequencies are restricted to words which have no case marking, hence which are ending in possessives.

```{r mA, echo = T, results = 'hide'}
person <- factor(c("1sg", "1pl", "2sg", "2pl", "3sg", "3pl"), levels = c("1sg", "1pl", "2sg", "2pl", "3sg", "3pl"))

freqs_ma <- c(91125,78437,14817,58381,1327200,319990)

ma_nom_df <- tibble(person, freqs_ma)

ma_nom_plot <- 
  ma_nom_df %>%
  mutate(props = freqs_ma/sum(freqs_ma)) %>%
  ggplot(aes(person, props, fill = person)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks=seq(0.1, 0.7, by=0.2), labels=scales::percent) +
  ggtitle("Relative Frequencies of Poss. Suffixes in NOM -mA clauses") +
  ylab("Relative Frequency") +
  xlab("Person-Number") +
  guides(fill="none")

ma_nom_plot
```